Fast and optimal centroiding of faint stars

### 2015-12-08 ms/figure.tex
edited the discussions section

An optimal stellar centroiding algorithm must saturate --or come close to saturating-- the fundamental Cram\'{e}r-Rao lower 
bound. That is, in all ranges of background noise level, size, radial light profile,
and shape, it must preserve information about the centroids of stars. In practice however,
this is only achievable when we have a reasonably good estimate of the PSF. Since we do not always 
know the exact PSF profile, we must make use of approximate centroiding algorithms. In this work, we
studied how close we get to saturating the CRLB with an approximate method acting on relatively low signal-to-noise ratio,
unsaturated stars.

@@

Moreover, it is important to note that the PSF fitting method can be made faster by only keeping the 
term proportional to the dot product of the PSF model and the image in $\chi^2$. This is because 
the other terms in $\chi^{2}$ are not sensitive to the position of the centroid. However, this only allows
us to vary the poition of centroid, and not the flux, while fitting the PSF model
to the star.

Modifying the $\chi^2$ such that it contains the dot product of the PSF and image, 
provides a nice interpretation for the PSF fitting method. It corresponds to searching for the 
centroid in a very high resolution grid which the dot product of the image and the PSF model is
sensitive to. On the other hand, one can think of the initial smoothing step in the default polynomial method 
as upsampling of the image around the center by convolving it with an approximate Gaussian PSF. 
The default polynomial centroiding is different from PSF-fitting method in that, when there is mismatch between the
widths of the smoothing kernel and the that of the PSF, we loose some information by employing a 3$\times$3 polynomial fitting.
When we have the advantage of knowing the PSF, this issue can be resolved by employing the matched-filter polynomial method.

Having a reasonable PSF model always helps us obtain more reliable centroid estimates, but
over a certain range of low signal-to-noise ratios and PSF sizes, one can achieve sensibly
 accurate results by employing a simple 3$\times$3 method after smooting the image with
 a Gaussian kernel of a fixed width, and without making any assumption about the PSF model at the positions of stars.

@@

This work was partially supported by the NSF (grants IIS-1124794 and AST-1517237), NASA (grant NNX12AI50G), 
and the Moore-Sloan Data Science Environment at NYU. We thank Jo Bovy for useful discussions.

>>>>

An optimal stellar centroiding algorithm must saturate---or come close to saturating---the fundamental Cram\'{e}r-Rao lower 
bound. That is, in all ranges of background noise level, size, radial light profile,
and shape, it must preserve information about the centroids of stars. In practice however,
this is only achievable when we have a reasonably good estimate of the PSF. Since we do not always 
know the exact PSF profile, we must make use of approximate centroiding algorithms. In this work, we
studied how close we get to saturating the CRLB with an approximate method acting on relatively low signal-to-noise ratio,
unsaturated stars.

@@

Moreover, it is important to note that the PSF fitting method can be made faster by only keeping the 
term proportional to the dot product of the PSF model and the image in $\chi^2$. This is because 
the other terms in $\chi^{2}$ are not sensitive to the position of the centroid. However, this only allows
us to vary the poition of centroid, and not the flux, while fitting the PSF model
to the star. This is the advantage of a matched filter approach.

Modifying the $\chi^2$ such that it only contains the dot product of the PSF and image, 
provides a nice interpretation. Finding a centroid coordinate that maximizes the dot product 
of the PSF and the star image is equivalent to finding the peak of the 
correlation of the PSF and the image. Therefore optimizing the modified $\chi^2$ is equivalent to finding the 
location of the peak of the matched filter.

On the other hand, one can think of the initial smoothing step in the default polynomial method 
as upsampling of the image around the center by correlating it with an approximate Gaussian PSF. 
The default polynomial centroiding is different from PSF-fitting method in that, when there is mismatch between the
widths of the smoothing kernel and the that of the PSF, we loose some information by employing a 3$\times$3 polynomial fitting.
When we have the advantage of knowing the PSF, this issue can be resolved by employing the matched-filter polynomial method.

Having a reasonable PSF model always helps us obtain more reliable centroid estimates, but
over a certain range of low signal-to-noise ratios and PSF sizes, one can achieve sensibly 
accurate results by employing a simple 3$\times$3 method after smoothing the image with 
a Gaussian kernel of a fixed width, and without making any assumption about the PSF model 
at the positions of stars.

@@

This work was partially supported by the NSF (grants IIS-1124794 and AST-1517237), NASA (grant NNX12AI50G), 
and the Moore-Sloan Data Science Environment at NYU. We thank Jo Bovy and Dustin Lang for useful discussions.
